---
title: "EPI590R Final Project"
author: "Sarah Whiteside"
execute:
  echo: false
format: html
---
## EPI 590R Final Project

## Central Park Squirrels Analysis

**Installation of packages**
This step installs packages that are not already installed on your computer neccessary for the following code to work. 
```{r}
#| messages: false
#| echo: false
#| warning: false
#| results: hide
packages_to_install <- c("tidytuesdayR", "gtsummary", "tidyverse", "here", "readr")
#install.packages(packages_to_install)
lapply(packages_to_install, require, character.only=TRUE)
```

## Reading in the dataset
```{r here} 
#| message: false
#| echo: false
#| warning: false
#| results: hide
here::here()
# my project directory is "/Users/swhiteside/Final Project EPI590R"
getwd()
# my working directory is "/Users/swhiteside/Final Project EPI590R"
squirrel_data <- read_csv(here("squirrel_data.csv"))
```

# *Central Park Squirrel Census*

## Squirrel data! The data this week comes from the 2018 Central Park Squirrel Census.

**Data Descriptions:** The Squirrel Census is a multimedia science, design, and storytelling project focusing on the Eastern gray (Sciurus carolinensis). They count squirrels and present their findings to the public. The dataset contains squirrel data for each of the 3,023 sightings, including location coordinates, age, primary and secondary fur color, elevation, activities, communications, and interactions between squirrels and with humans.

**Cleaning the dataset**
```{r cleaning}
squirrel_data$Age[squirrel_data$Age == "?"] <- NA
```
This step is neccessary as some of the missing values for Age were coded as NA and others as question marsk (?). This homogenizes the missing values as NA. 

**Table 1. Descriptive Statistic**
```{r}
#| label: tbl-descriptive
#| tbl-cap: "Table 1. Descriptive Statistics" 
#| message: false
#| warning: false
#| results: hide

descriptive_table <- tbl_summary(
	squirrel_data,
	by = "Primary Fur Color",
	include = c("Age", "Shift", "Running", "Chasing", "Climbing", "Foraging"),
	missing_text = "Missing") |> 
		add_p(test = list (all_continuous() ~ "t.test",
											 all_categorical() ~ "chisq.test")) |> 
		add_overall(last=TRUE) |> 
		bold_labels() |> 
		modify_footnote(update = everything() ~ NA) |> 
		modify_header(label = "**Variable**", p.value = "**P**") |> 
    modify_caption(caption = "Table 1. Descriptive Statistics")
descriptive_table 


inline_text(descriptive_table, variable="Age", column = "Black", level = "Juvenile") #8 (8%)
inline_text(descriptive_table, variable="Age", column = "Gray", level = "Juvenile") #256 (11%)
inline_text(descriptive_table, variable="Age", column = "Cinnamon", level = "Juvenile") # 58 (15%)
inline_text(descriptive_table, variable="Age", column = "Black", level = "Adult") #92 (92%)
inline_text(descriptive_table, variable="Age", column = "Gray", level = "Adult") #2125 (89%)
inline_text(descriptive_table, variable="Age", column = "Cinnamon", level = "Adult") #326 (85%)

# Using here to save table as a csv 
descriptive_table_df <- as.data.frame(descriptive_table)
write_csv(descriptive_table_df, here("descriptive_table.csv"))
```


@tbl-descriptive describes the distribution of certain characteristics of the squirrels in Central Park. One of these characteristics is Age, which is displayed by primary fur color.  

For the juvenile age category, there were `r inline_text(descriptive_table, variable = "Age", column = "Black", level = "Juvenile")` Black juvenile squirrels and `r inline_text(descriptive_table, variable = "Age", column = "Black", level = "Adult")` adult Black squirrels.  

There were `r inline_text(descriptive_table, variable = "Age", column = "Gray", level = "Juvenile")` juvenile Gray squirrels and `r inline_text(descriptive_table, variable = "Age", column = "Gray", level = "Adult")` adult Gray squirrels.  

Lastly, there were `r inline_text(descriptive_table, variable = "Age", column = "Cinnamon", level = "Juvenile")` juvenile Cinnamon squirrels and `r inline_text(descriptive_table, variable = "Age", column = "Cinnamon", level = "Adult")` adult Cinnamon squirrels.

**Table 2. Regression Table**
```{r} 
#| label: tbl-regression
#| tbl-cap: "Table 2. Binomial Logistic Regression"
#| message: false
#| warning: false

# Recode binary predictors
binary_func <- function(data, cols) {
  data[cols] <- lapply(data[cols], function(x) ifelse(x, 1, 0))
  data
}
squirrel_data <- binary_func(squirrel_data, c("Tail flags", "Tail twitches"))

# Recode outcome as factor
squirrel_data$Age <- factor(
  ifelse(squirrel_data$Age == "Adult", 1,
  ifelse(squirrel_data$Age == "Juvenile", 0, NA)),
  levels = c(0, 1),
  labels = c("Juvenile", "Adult"))

# Run regression
table2 <- tbl_uvregression(
  data = squirrel_data,
  y = Age,
  include = c(`Tail flags`, `Tail twitches`),
  method = glm,
  method.args = list(family = binomial()),
  exponentiate = TRUE
) |> 
  modify_header(label ~ "**Predictors**") |>
  modify_caption("**Table 2. Binomial Logistic Regression - Tail Flags and Twitches as Predictors of Squirrel Age**")

table2
```

@tbl-regression is a binomial logistic regression with squirrel age as the outcome and tail flags and tail twitches as predictors.


## **Figure**
```{r}
#| label: fig-barplot
#| fig-cap: "Barplot of Primary Fur Color"
#| message: false
#| results: hide

# Creating figure
barplot(table(squirrel_data$`Primary Fur Color`),
        col = c("black", "firebrick", "gray"),
        main = "Distribution of Primary Fur Color",
        ylab = "Count",
        xlab = "Primary Fur Color")

# Calculating how many squirrels in each primary fur color
table(squirrel_data$`Primary Fur Color`)["Black"]
table(squirrel_data$`Primary Fur Color`)["Cinnamon"]
table(squirrel_data$`Primary Fur Color`)["Gray"]
```

@fig-barplot contains a barplot of primary fur color by squirrels spotted inside Central Park. There were `r table(squirrel_data[["Primary Fur Color"]])["Black"]` squirrels with Black as the primary fur color, `r table(squirrel_data[["Primary Fur Color"]])["Cinnamon"]` squirrels with Cinnamon as the primary color, and `r table(squirrel_data[["Primary Fur Color"]])["Gray"]` squirrels with Gray as the primary color.

